Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-06-28 paper_jasa.tex
model fitting discussion

The model in Equation \ref{eqn:mixed} provides the subject and lineup specific probability of successful evaluation as 
\begin{equation} \label{eqn:mixed_power}
\hat p_{\ell i} = \frac{exp(\hat g_{\ell i})}{1+ exp(\hat g_{\ell i})}
\end{equation}
where $\hat g_{\ell i} = W_{\ell i} \hat {\delta} +  Z_{\ell i}  \hat {\tau}_{\ell i}$.

@@

The model in Equation \ref{eqn:mixed} is fit using $E$ as the only fixed effect covariate without intercept. we set offset = $log(0.05/0.95)$ so that the estimated power has a fixed lower limit at 0.05 (Type-I error) when $E=0$. Different subjects are expected to have different skill levels in evaluating lineups. To account for subject specific variation we use subject as random effect. We also use an interaction between subject and effect $E$ to allow for subject specific random slope. This leads to setting  $W_{\ell i} = E_{\ell i}$, the effect of lineup $\ell$ that subject $i$ has evaluated in Equation \ref{eqn:mixed}.  While for experiment 1 and 2 we don't allow intercept but offset, for experiment 3 an intercept is estimated to allow the power curves to go above 0.05 when $E=0$. This is because forcing power to be fixed at 0.05 for $E=0$ is not justified by the experiment 3 data (Figure \ref{fig:power_loess_effect}).

@@

\begin{table}[hbtp]
\caption{Parameter estimates of model \ref{eqn:mixed}.}
\begin{center}
\begin{tabular}{ccc}
  \hline
 &  \multicolumn{2}{c} {Estimates}  \\
 \cline{2-3}
 Experiment & Effect & inter subject correlation\\
  \hline
  1 & 0.39 &  \\ 
  2 & 1.22 &  \\ 
  3 &  & \\ 
   \hline
\end{tabular}
\end{center}
\label{tbl:model_par}
\end{table}

>>>>

The model in Equation \ref{eqn:mixed} provides the subject and lineup specific probability of successful evaluation as 
\begin{equation} \label{eqn:mixed_power}
\hat p_{\ell i} =  g^{-1}(W_{\ell i} \hat {\delta} +  Z_{\ell i}  \hat {\tau}_{\ell i}).
\end{equation}

@@

The model in Equation \ref{eqn:mixed} is fit using $E$ as the only fixed effect covariate without intercept. we set offset = $log(0.05/0.95)$ so that the estimated power has a fixed lower limit at 0.05 (Type-I error) when $E=0$. Different subjects are expected to have different skill levels in evaluating lineups. To account for subject specific variation we use subject as random effect. We do not include random intercept, instead, we use an interaction between subject and effect $E$ to allow for subject specific random slope. This leads to setting  $W_{\ell i} = E_{\ell i}$, the effect of lineup $\ell$ that subject $i$ has evaluated in Equation \ref{eqn:mixed}.  While for experiment 1 and 2 we don't allow intercept but offset, for experiment 3 an intercept is estimated to allow the power curves to go above 0.05 when $E=0$. This is because forcing power to be fixed at 0.05 for $E=0$ is not justified by the experiment 3 data (Figure \ref{fig:power_loess_effect}).

@@

\begin{table}[hbtp]
\caption{Parameter estimates of model in Equation \ref{eqn:mixed}.}
\begin{center}
\begin{tabular}{cccc}
  \hline
 &  \multicolumn{2}{c} {Fixed effect}  & Random effect\\
 \cline{2-3}
 Experiment & Estimate & Std. error & Variance\\
  \hline
  1 & 0.39 & 0.0093 & 0.0077 \\ 
  2 & 1.22 &  0.0196 &  0.0455 \\ 
  3 &  & & \\ 
   \hline
\end{tabular}
\end{center}
\label{tbl:model_par}
\end{table}
